---
output:
  html_notebook: default
  html_document: default
---
## Lab #2 -- Temporal Processes


## Discrete event history analysis -- Repeated events
```{r message=FALSE, warning=FALSE}
# Install packages
library(plm)
library(plyr)
library(dplyr)
library(tidyverse)
library(stargazer)
library(ggplot2)

# Import data file and generate a subset
vars = c("marital", "age", "year","educ", "idnum", "panelwave", "sex")
sub = data.table::fread('/Users/hengyuai/Desktop/TS/Lab2/panel-for-R.csv',
                        sep = ",",
                        select = vars)
```

### (a) State what your "failure" variable is and how you expect your independent variables to affect it.
```{r}
# My failure variable is 'divorce', if the  and independent vairable is "age", "year", "sex" and "educ"
# Recode the variable 'marital'
sub[sub$marital < 6]
sub$divorce = ifelse(sub$marital == 3, 1, 0) # it's 1 if the respondent divorced, 0 if not.
```

```{r}
# Find all the people from Wave 1 that I want to remove: those already divorced 
sub$drop = ifelse(sub$divorce == 1 & sub$panelwave == 1, 1, 0)

# Because I have 3 rows per person, I need to drop all three of those rows at once, so I do that with merging and subsetting them away
sub = merge(sub, aggregate(drop~idnum, data = sub, FUN = mean), by="idnum", suffixes=c("", ".all"))

subma = subset (sub, sub$drop.all==0)

# If using the code below, we can only drop the first year data of people who divorce at the panelwave1
# sub$divbegin1 = ifelse(sub$divorce == 1 & sub$panelwave == 1, 1, 0)
# sub = sub[sub$divbegin1 == 0]
```

```{r}
# I want to see when people “die” so I look to see if it happens in Wave 2 or in Wave 3, or both
subma$ytwo = ifelse(subma$divorce == 1 & subma$panelwave == 2, 2, 0)
subma = merge(subma, aggregate(ytwo~idnum, data = subma, FUN = max), by="idnum", suffixes=c("",".two"))
subma$ythree  = ifelse(subma$divorce == 1 & subma$panelwave == 3, 3, 0)
subma = merge(subma, aggregate(ythree~idnum, data = subma, FUN = max), by="idnum", suffixes = c("",".three"))
```

```{r}
# Now, I see what their pattern of death was across time

# subma$combo <- do.call(paste, c(subma[c("ytwo.two", "ythree.three")], sep = ""))
subma = subma %>% unite(combo, ytwo.two, ythree.three)
table(subma$combo)
```

```{r}
table(subma$divorce, subma$combo)
```

```{r}
# Make a lagged divorce indicator
subma$di = ifelse(subma$combo == "2_0" & subma$panelwave == 3 | subma$combo == "2_3" & subma$panelwave == 3, 1, 0)

# table(subma$dlag)
table(subma$di)
```

```{r}
# 63
subma$nyear = subma$year - 2006
logit_lag = glm(divorce ~ nyear, subma, family="binomial", subset = subma$year>2006)
summary(logit_lag)
```

More obs. (2319 vs. 2355) people can contribute 2 events over 2 Waves now.
For Wave 3, relative to Wave 2, the logit increases by 0.159 of divorcing, when divorce can happen over and over again.

```{r}
logit_lag2 = glm(divorce ~ as.factor(nyear) + di, subma, family="binomial", subset = subma$year>2006)
summary(logit_lag2)
```

With an earlier time being already divorced, the logit increases by 4.92 of being divorced now, net of time.

```{r}
logit_lag3 = glm(divorce ~ as.factor(nyear) + di + sex + educ + age, subma, family="binomial", subset = subma$year>2006)
summary(logit_lag3)
```

For each year more education someone has, their logit of divorcing decreases by -0.0483, when divorce can happen over and over again and net of the lag of getting divorced the time before and other factors


## Multiple outcomes
```{r}
subma$marriage[subma$marital==3] <- 3 #divorced
subma$marriage[subma$marital==4] <- 2 #seperated
subma$marriage[subma$marital==1] <- 1 #married

library(mlogit)
subma2 = mlogit.data(subma, varying=NULL, choice="marriage", shape="wide")
```

```{r}
# With multiple outcomes possible
mari1 = mlogit(marriage ~ 1 | panelwave, data=subma2, reflevel="1", subset =
subma2$year>2006)
summary(mari1)
```

Relative to Wave 2, Wave 3 lowers the logit on getting seperated by -0.00820, compared to getting married
Relative to Wave 2, Wave 3 increases the logit on being divorced by 0.3099+, compared to getting married

```{r}
# With multiple outcomes possible II
mari2 = mlogit(marriage ~ 1 | panelwave + sex + educ + age, data=subma2,
reflevel="1", subset = subma2$year>2006)
summary(mari2)
```

Education predicts being seperated (b=-0.1205) nearly twice higher as much as getting divorced employment (b=-0.0774), relative to unemployment, net of other factors

## 2. Continuous survival analysis
```{r}
# The Kaplan-Meier survival estimates
# A nonparametric estimate of the survivor function, S(t), which is the probability of survival past time t
library(survival)
plot(survfit(Surv(subma$panelwave, subma$divorce) ~ 1, subset=subma$year>2006))
```


```{r}
# 1a. One-time events
# Cox (multivariable)
subma$time = 1

coxph1 <- coxph(Surv(subma$time,subma$divorce)~subma$age + subma$educ +
subma$sex, method="efron", subset = subma$year>2006)
summary(coxph1)
```

For each year of age, the log hazard rate (i.e., the instantaneous rate of change) increase by 0.9896, net of all factors.
Haz. ratio of educ = exp(1.0489). For each additional year of education, the rate of becoming divoced increases by 4.9% (e.g., 1.0489-1=.0489), net of all other factors.

```{r}
coxph2 = coxph(Surv(subma$time,subma$divorce)~subma$age + subma$educ +
subma$sex, method="breslow", subset = subma$year>2006)
stargazer(coxph1, coxph2, type="text", column.labels = c("Efron", "Breslow"))
```

```{r}
# The all ages model vs. the restricted one
coxph3 = coxph(Surv(subma$time,subma$divorce)~subma$age + subma$educ +
subma$sex, method="efron", subset = subma$year>2006 & subma$age>22 &
subma$age<60)
stargazer(coxph1, coxph3, type="text", column.labels = c("All Ages",
"Age 23 to Age 59"))
```

```{r}
# The Schoenfeld test -- in R
coxph1 = coxph(Surv(subma$time,subma$divorce)~subma$age + subma$educ +
subma$sex, method="efron", subset = subma$year>2006, na.action=na.exclude)
zph.coxph1 = cox.zph(coxph1, transform = 'log')
print(zph.coxph1)
```

```{r}
# The survival curves of people with and without high education level
library(prodlim)
subma$hiedu = ifelse(subma$educ >= 15, 1, 0)
km1 <- prodlim(Hist(panelwave,divorce) ~ hiedu,data=subma)
par(mar=c(7,7,5,5), mgp=c(3,1,0))
plot(km1, legend.x="bottomleft")
```

```{r}
mean(subma$divorce[subma$year==2008 & subma$hiedu==1], na.rm=T)
```
```{r}
mean(subma$divorce[subma$year==2008 & subma$hiedu==0], na.rm=T)
```
```{r}
mean(subma$divorce[subma$year==2010 & subma$hiedu==1], na.rm=T)
```
```{r}
mean(subma$divorce[subma$year==2010 & subma$hiedu==0], na.rm=T)
```
People with higher eductaion are “failing” (into divorce) at higher and faster rates than people without higher education in 2008. But in 2010, people with higher eductaion are “failing” (into divorce) at lower and slower rates than people without higher education.









