---
output:
  html_notebook: default
  html_document: default
---
## Lab #2 -- Temporal Processes

1. subpt$combo <- do.call(paste, c(subpt[c("ytwo.two", "ythree.three")], sep = ""))doesn't work
2. logit1 = glm(divorce~as.factor(nyear), subma, family = "binomial", subset = subma$year > 2006) why specify >2006. have already removed all divorced in panelwave1
3. why cluster? why exp algorithm? exp(coef(logit))?
4. divorcing logit. why it's different between male and female
5. ytwo.two replace all NAs by 0, bad impact?
6. logit_lag nyear and logit uses as.factor(nyear). if change one of them, same result. 
7. difference between cox and Kaplan-Meier
8. 83 84
9. Schoenfeld test??


## Survival Analysis 
### Run a multiple variable survival analysis. You can perform the survival analysis either using discrete-time methods (i.e., event history analysis) or you can use Cox proportional hazards methods, either one is fine.
### Explain how you determined the #risk window" (due to right truncation and left-censoring) and who is eligible for failure over the time you are studying.  
### Explain whether the results were consistent with your expectations, and do that by interpreting the coefficients from the models, model fit, and so on.

##Discrete event history analysis -- One-time events
```{r message=FALSE, warning=FALSE}
# Install packages
library(plm)
library(plyr)
library(dplyr)
library(tidyverse)
library(stargazer)
library(ggplot2)

# Import data file and generate a subset
vars = c("marital", "age", "year","educ", "idnum", "panelwave", "sex")
sub = data.table::fread('/Users/hengyuai/Desktop/TS/Lab2/panel-for-R.csv',
                        sep = ",",
                        select = vars)
```

### (a) State what your "failure" variable is and how you expect your independent variables to affect it.
```{r}
# My failure variable is 'divorce', if the  and independent vairable is "age", "year", "sex" and "educ"
# Recode the variable 'marital'
sub[sub$marital < 6]
sub$divorce = ifelse(sub$marital == 3, 1, 0) # it's 1 if the respondent divorced, 0 if not.
```

```{r}
# Find all the people from Wave 1 that I want to remove: those already divorced 
sub$drop = ifelse(sub$divorce == 1 & sub$panelwave == 1, 1, 0)

# Because I have 3 rows per person, I need to drop all three of those rows at once, so I do that with merging and subsetting them away
sub = merge(sub, aggregate(drop~idnum, data = sub, FUN = mean), by="idnum", suffixes=c("", ".all"))

subma = subset (sub, sub$drop.all==0)

# If using the code below, we can only drop the first year data of people who divorce at the panelwave1
# sub$divbegin1 = ifelse(sub$divorce == 1 & sub$panelwave == 1, 1, 0)
# sub = sub[sub$divbegin1 == 0]
```

```{r}
# I want to see when people “die” so I look to see if it happens in Wave 2 or in Wave 3, or both
subma$ytwo = ifelse(subma$divorce == 1 & subma$panelwave == 2, 2, 0)
subma = merge(subma, aggregate(ytwo~idnum, data = subma, FUN = max), by="idnum", suffixes=c("",".two"))
subma$ythree  = ifelse(subma$divorce == 1 & subma$panelwave == 3, 3, 0)
subma = merge(subma, aggregate(ythree~idnum, data = subma, FUN = max), by="idnum", suffixes = c("",".three"))
```

```{r}
# Now, I see what their pattern of death was across time

# subma$combo <- do.call(paste, c(subma[c("ytwo.two", "ythree.three")], sep = ""))
subma = subma %>% unite(combo, ytwo.two, ythree.three)
table(subma$combo)
```

```{r}
table(subma$divorce, subma$combo)
```

```{r}
# Making missing in panelwave3 anybody who already died in panelwave2
subma$divorce[subma$combo == "2_0" & subma$panelwave == 3] = NA
subma$divorce[subma$combo == "2_3" & subma$panelwave == 3] = NA

table(subma$divorce, subma$combo)
```

```{r}
# This model just has the wave variable to see if the underlying hazard rate changes over time, >2006
subma$nyear = subma$year - 2006

# logit = glm(divorce~as.factor(nyear), subma, family = "binomial")
# summary(logit)

#44
logit = glm(divorce~as.factor(nyear), subma, family = "binomial", subset = subma$year > 2006)
summary(logit)
```
Wave 3 lowers the logit on divorcing by -0.411, relative to Wave 2. The coefficient is not statistically significantly.


```{r}
# clustering
exp(coef(logit))
```
Wave 3 lowers the odds of finding full-time employment by 34% (=(0.66-1)*100%), relative to Wave 2; there isn't a statistically significantly lower risk of finding full-time employment in Wave 3.
Wave 3 lowers the hazard ratio of divorcing by about 34%.

```{r}
# Adding some actual variables
logit2 = glm(divorce~as.factor(nyear) + age + educ + as.factor(sex), subma, family="binomial", subset=subma$year > 2006)
summary(logit2)
```
For each year older someone becomes, their logit of divorcing decreases by -0.0124, net of other factors. 
For each year more education someone has, their logit of divorcing decreases by -0.0442, net of other factors.

```{r}
# what about more "younger" people?
logit3 = glm(divorce~as.factor(nyear) + age + educ + sex, subma, family="binomial", subset=subma$year > 2006 & subma$age > 22 & subma$age < 59)
summary(logit3)
```

The coefficients of age and educ both got stronger.

```{r}
# Consider interacting some of the Xs with the time variables
logit4 = glm(divorce~as.factor(nyear) * as.factor(sex) + age + educ, subma, family="binomial", subset=subma$year > 2006 & subma$age > 22 & subma$age < 59)
summary(logit4)
```
At baseline, women have a 0.1142 logit of divorcing, compared to men … but for Wave 3, relative to Wave 2, they have a logit 0.485 higher of divorcing, compared to men, net of other factors











