---
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

## Option 3. Conduct an age-period-cohort analysis. Make plots of the different dimensions. Develop a parsimonious model of what you think is going on. Explain your results.

Question：pattern, parsimonious，summary collinearity, rho

## install packages and open the data file
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(QMSS)
library(ggplot2)
library(plyr)
library("rms") # to do splines
library(Epi)
vars = c("chldidel", "age", "year", "cohort", "sex")
sub = data.table::fread("/Users/hengyuai/Desktop/TS/Lab1/trends-gss.csv",
                        sep = ",",
                        select = vars)

sub = na.omit(sub)
```


## Break year, age and cohort each into 10 groups
```{r}
sub = mutate(sub, 
             year.cut = cut(year, breaks = 10, labels = F, right = F),
             age.cut = cut(age, breaks = 10, labels = F, right = F),
             cohort.cut = cut(cohort, breaks = 10, labels = F, right = F))
```


## Tabulate age-profiles for each cohort
```{r}
tab.age.cohort = stat.table(index = list("Age" = age.cut, "Cohort" = cohort.cut),
                        contents = mean(chldidel),
                        data = sub,
                        margins = T)


print(tab.age.cohort, digits = 3)
```


## Graph age-profile for each cohort
```{r}
age.cohort = ddply(sub, c("age.cut", "cohort.cut"), summarise, CHLDIDEL = mean(chldidel))
graph.age.cohort = ggplot(age.cohort, aes(x = age.cut, y = CHLDIDEL, group = cohort.cut, color = factor(cohort.cut)))
graph.age.cohort = graph.age.cohort + geom_point(size = 3) + geom_line()
graph.age.cohort
```


## Tabulate cohort chldidel scores over time
```{r}
tab.year.cohort = stat.table(index = list("Year" = year.cut, "Cohort" = cohort.cut),
                             contents = mean(chldidel),
                             data = sub,
                             margins = T)
print(tab.year.cohort, digits = 3)
```


## Graph chldidel overtime
```{r}
year.cohort = ddply(sub, c("year.cut", "cohort.cut"), summarise, CHLDIDEL = mean(chldidel))
graph.year.cohort = ggplot(year.cohort, aes(x = year.cut, y = CHLDIDEL, group = cohort.cut, color = factor(cohort.cut)))
graph.year.cohort = graph.year.cohort + geom_point(size = 3) + geom_line()
graph.year.cohort
```


## Tabulate age-profiles by survey years
```{r}
tab.age.year = stat.table(index = list("Year" = year.cut, "AGE" = age.cut),
                             contents = mean(chldidel),
                             data = sub,
                             margins = T)
print(tab.age.year, digits = 3)
```


## Graph age-profiles by survey years
```{r}
age.year = ddply(sub, c("age.cut", "year.cut"), summarise, CHLDIDEL = mean(chldidel))
graph.age.year = ggplot(age.year, aes(x = age.cut, y = CHLDIDEL, group = year.cut, color = factor(year.cut)))
graph.age.year = graph.age.year + geom_point(size = 3) + geom_line()
graph.age.year
```

It seems that there is an age pattern in the data.
A clear cohort differences in mean chldidel.
Strong period effects, it looks for all cohorts and ages in the recent period, they raise their chldidel in the recent decades.


## Test collinearity
```{r}
summary(lm(chldidel ~ cohort + age + year, data = sub))
```


## Model 1: An everything model
```{r}
lm.chldidel = lm(chldidel ~ factor(age.cut) + factor(year.cut) + factor(cohort.cut), data = sub)
summary(lm.chldidel)
```


## Which years correspond to the year cuts?
```{r}
ddply(sub, "year.cut", summarise, min = min(year), max = max(year))
```


## Model2: Try to reduce this whole model just to period effects? … like this:
```{r}
sub$late = ifelse(sub$year > 1994, 1, 0)
lm.chldidel.apc = lm(chldidel ~ age + late + factor(cohort.cut), data = sub)
summary(lm.chldidel.apc)
```


## Compare R^2
```{r}
c(Model1 = summary(lm.chldidel)$adj, Model2 = summary(lm.chldidel.apc)$adj)
```