
## Conduct a difference-in-difference analysis.  Explain which group is treated and which is not.  Provide pre-treatment trends and generate a regression summarizing your results.  Explain what you found. 

===

关于pdf里的parallel trends assumption，并没有single women的vairable，如何确定就是parallel

## Open the data file
```{r, message=FALSE, echo=FALSE}
library(QMSS)
require(foreign)
sub = read.dta("/Users/hengyuai/Desktop/TS/Lab1/eitc.dta")

```

## Calculate the mean of different groups
```{r}
mean(sub[sub$children == 0, "work"])
```

```{r}
mean(sub[sub$children == 1, "work"])
```

```{r}
mean(sub[sub$children >= 1, "work"])
```

## recode
```{r}
sub$c.earn = sub$earn * sub$work #Make sure that people without work can't get any earnings
z = names(sub)
X = as.data.frame(sub$c.earn)
X[ ] = lapply(X, function(x){replace(x, x == 0, NA)})
sub = cbind(sub, X)
sub$c.earn = NULL

sub$post93 = as.numeric(sub$year >= 1994)
sub$anykids = as.numeric(sub$children > 0)

# Take average value of 'work' by year, conditional on highedu
minfo = aggregate(sub$work, list(sub$year,sub$anykids == 1), mean)
minfo
```

## Look at the table of means by group, over time
```{r}
# rename column headings (variables)
names(minfo) = c("YR","Treatment","LFPR") #LFPR means "labor force participation rate"
# Attach a new column with labels
minfo$Group[1:6] = "Single women, no children"
minfo$Group[7:12] = "Single women, children"
minfo
```

## Make a graph
```{r}
library(ggplot2)
qplot(YR, LFPR, data = minfo, geom = c("point", "line"), color = Group, xlab = "Year", ylab = "Labor Force Participation Rate")
```

## Just looking at differences in means between single women without kids vs. single women with any kids
```{r}
a = colMeans(subset(sub, post93 == 0 & anykids == 0, select = work))
b = colMeans(subset(sub, post93 == 1 & anykids == 0, select = work))
c = colMeans(subset(sub, post93 == 0 & anykids == 1, select = work))
d = colMeans(subset(sub, post93 == 1 & anykids == 1, select = work))

(d-b)-(c-a)
```

## In regression term
```{r}
reg1 = lm(work ~ post93 + anykids + post93 * anykids, data = sub)
summary(reg1)
```
post93:anykids的cofficient就是(d-b)-(c-a)的difference in differences

## Add covariates
```{r}
options(scipen = 999)
reg2 = lm(work ~ post93 + anykids + post93 * anykids + nonwhite + age + I(age^2) + ed + finc +I(finc-earn), data = sub)
summary(reg2)
```

## Try a different year to see if I can induce a stat. sig. result and check if choosing the year 94 is correct
```{r}
# sub set the data, including only years before 1994.
subbefore = sub[sub$year <= 1993, ]

# 假设可能的treatment在92年。Create a new "after treatment" dummy variable and interaction term
subbefore$post91 = as.numeric(subbefore$year >= 1992)
reg3 = lm(work ~ anykids  + post91 + post91 * anykids, data = subbefore)
summary(reg3)
```
The coefficient is not significant.




